---
layout: post
title: Number of COVID-19 deaths in V4 countries
subtitle: ggplot linechart example
output:
  md_document:
    variant: gfm
    preserve_yaml: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_posts") })
categories: [R stat]
tags: [ggplot, linechart, COVID-19]
lang: en
---


I would like to highlight here the charting capabilities of ggplot, that are based on different layers containing geoms (lines, points, bars, etc). In the very basic ggplot declaration there should be an "aesthetic mapping" that describes how variables in the data are mapped to geoms. However, additional mappings can be defined in each layers.

We shall create a line chart, which shows the cumulative daily COVID-19 deaths in Visegrad countries, that are Czechia, Hungary, Poland and Slovakia with R's ggplot package. 

```{r include = FALSE}
base_dir <- "~/projects/bj/_source" # i.e. where the jekyll blog is on the hard drive.
base_url <- "/" # keep as is
fig_path <- "img/" # customize to heart's content, I 'spose.

knitr::opts_knit$set(base.dir = base_dir, base.url = base_url)
knitr::opts_chunk$set(fig.path = fig_path,
                      cache.path = '../cache/',
                      message=FALSE, warning=FALSE,
                      cache = TRUE) 
```
```{r echo = FALSE, eval = TRUE, results = "asis"}
cat('{% include image.html url="/images/brent-rambo.gif" caption="Jekyll rules the world. Your Jekyll friend, Steve" width=410 align="right" %} ')
```

```{r}
# setwd("/home/fabian/projects/2020/linechart")
library(tidyverse)
library(extrafont)
# download and read csv file
uri <- c("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
uri
df <- read_csv(uri)

# some data wrangling - cf. my previous post
df <- df %>% select(-c("Province/State", "Lat", "Long"))
df <- df %>% rename(c("country" = "Country/Region"))
df <- pivot_longer(df, cols = ends_with("20"),
    names_to = "date", values_to = "deaths")
df <- df %>% filter(country == c("Hungary", "Slovakia", "Poland", "Czechia"))
df$date <- as.Date(df$date, "%m/%d/%y")
df <- arrange(df, date, country)
tail(df)

# create the plot
mycolors <- c("#FF9700", "black", "red", "blue") # define 4 colours for the countries 
p1 <- ggplot(df, aes(date, deaths, group = country))
p1 <- p1 + geom_line(aes(color = country, linetype = country), size = 2) # note that color and line mappings ar inside aes()
p1 <- p1 + geom_point(aes(shape = country, color = country), size = 4) # again: aes () contains mappings
p1 <- p1 + scale_color_manual(values = mycolors) # manual color setting instead of default
p1 <- p1 + scale_linetype_manual(values =
     c("solid", "twodash", "solid", "longdash")) # custom linetype
p1 <- p1 + theme(legend.position = "top", legend.title = element_blank()) # move legend to top & supress legend title
# p1 <- p1 + theme(text = element_text(family = "Ubuntu Mono", size = 12)) # custom TTF font with extrafonts pkg
p1 <- p1 + labs(title = "Number of COVID deaths in V4 countries",
              subtitle = "Cumulative numbers",
              caption = "Data source: JHU")
p1 <- p1 + xlim(x = c(Sys.Date() - 60, NA)) # x scale range: previous 60 days 
p1
```



That's all. [Download the Rscript file]().

